---
title: "SPOTVignetteNutshell"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SPOTVignetteNutshell}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitrSetup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r SPOT, include = FALSE, eval = FALSE}
## install.packages("devtools")
## devtools::install_github("r-lib/devtools")
url <- "http://owos.gm.fh-koeln.de:8055/bartz/spot.git"
devtools::install_git(url = url)
```

* Package version SPOT should be larger than `2.3.0`.

```{r, setup}
library("SPOT")
packageVersion("SPOT")
```

## Introduction

* The performance of modern search heuristics relies crucially 
on their parameterizations---or, statistically speaking, on their factor settings. 
* Finding good parameter settings for an optimization algorithm will be referred to as *tuning*.
* The version of `SPOT` presented in this article is implemented in `R`.
* It can be used as an algorithm tuner or directly as an optimizer.
* This vignette is an abbreviated version of a the technical report "In a Nutshell-The Sequential Parameter Optimization Toolbox".
* It illustrates how `spot` can be called. 
* The extended version of this vignette is available on [arXiv](https://arxiv.org/abs/1712.04076) and will be updated regularly.

## Sequential Parameter Optimization Examples: How to Call SPOT

### Most simple example: Kriging + LHS + predicted  mean optimization (not expected improvement)

```{r, spotSimple}
res <- spot(,funSphere,c(-2,-3),c(1,2),control=list(funEvals=15))
res$xbest
```

### With expected improvement

```{r, spotEI}
res <- spot(,funSphere,c(-2,-3),c(1,2),
    control=list(funEvals=15,modelControl=list(target="ei")))
res$xbest
```

### With additional start point:

```{r, spotStart}
res <- spot(matrix(c(0.05,0.1),1,2),funSphere,c(-2,-3),c(1,2))
res$xbest
```

### Larger budget:

```{r, spotLarger}
res <- spot(,funSphere,c(-2,-3),c(1,2),
    control=list(funEvals=50))
res$xbest
```

### Use local optimization instead of LHS

```{r, spotLocal}
res <- spot(,funSphere,c(-2,-3),c(1,2),
   control=list(optimizer=optimLBFGSB))
 res$xbest
```

### Random Forest instead of Kriging

```{r, spotRf}
res <- spot(,funSphere,c(-2,-3),c(1,2),
     control=list(model=buildRandomForest))
res$xbest
```

### LM instead of Kriging

```{r, spotLM}
res <- spot(,funSphere,c(-2,-3),c(1,2),
     control=list(model=buildLM)) #lm as surrogate
res$xbest
```

### Bayesian Optimization

```{r, spotBO}
res <- spot(,funSphere,c(-2,-3),c(1,2),
     control=list(model=buildBO)) #BO as surrogate
res$xbest
```



### LM and local optimizer (which for this simple example is perfect)
```{r, spotLMOPT}
res <- spot(,funSphere,c(-2,-3),c(1,2),
   control=list(model=buildLM, optimizer=optimLBFGSB))
res$xbest
```

### Lasso and local optimizer NLOPTR

```{r, spotLasso}
res <- spot(,funSphere,c(-2,-3),c(1,2), 
   control=list(model=buildLasso, optimizer = optimNLOPTR))
res$xbest
```

### Kriging and local optimizer LBFGSB 

```{r, spotLBFGSB}
res <- spot(,funSphere,c(-2,-3),c(1,2), 
   control=list(model=buildKriging, optimizer = optimLBFGSB))
res$xbest
```

### Kriging and local optimizer NLOPTR 

```{r, spotNLOPTR}
res <- spot(,funSphere,c(-2,-3),c(1,2), 
     control=list(model=buildKriging, optimizer = optimNLOPTR))
res$xbest
```

### Or a different Kriging model:

```{r, spotKrigDace}
res <- spot(,funSphere,c(-2,-3),c(1,2),
 control=list(model=buildKrigingDACE, optimizer=optimLBFGSB))
res$xbest
```

### With noise: (this takes some time)

```{r, spotNoise}
# noisy objective
res1 <- spot(,function(x)funSphere(x)+rnorm(nrow(x)),c(-2,-3),c(1,2),
 		control=list(funEvals=40,noise=TRUE)) 
# noise with replicated evaluations
res2 <- spot(,function(x)funSphere(x)+rnorm(nrow(x)),c(-2,-3),c(1,2),
 		control=list(funEvals=40,noise=TRUE,replicates=2,
 		designControl=list(replicates=2))) 
# and with OCBA
res3 <- spot(,function(x)funSphere(x)+rnorm(nrow(x)),c(-2,-3),c(1,2),
 		control=list(funEvals=40,noise=TRUE,replicates=2,OCBA=TRUE,OCBABudget=1,
 		designControl=list(replicates=2))) 
# Check results with non-noisy function:
funSphere(res1$xbest)
funSphere(res2$xbest)
funSphere(res3$xbest)
```

### Random number seed handling

The following is for demonstration only, to be used for random number
seed handling in case of external noisy target functions.

```{r, randomNumberSeed1a}
res1a <- spot(,function(x,seed){set.seed(seed);funSphere(x)+rnorm(nrow(x))},
     c(-2,-3),c(1,2),control=list(funEvals=25,noise=TRUE,seedFun=1))
```

```{r, randomNumberSeed1b}
res1b <- spot(,function(x,seed){set.seed(seed);funSphere(x)+rnorm(nrow(x))},
     c(-2,-3),c(1,2),control=list(funEvals=25,noise=TRUE,seedFun=1))
```

```{r, randomNumberSeed2}
res2 <- spot(,function(x,seed){set.seed(seed);funSphere(x)+rnorm(nrow(x))},
     c(-2,-3),c(1,2),control=list(funEvals=25,noise=TRUE,seedFun=2))
```


```{r, printRes1a1b2}
sprintf("Should be equal: %f = %f. Should be different:  %f", res1a$ybest, res1b$ybest, res2$ybest)
```




### Handling factor variables

Note: factors should be coded as integer values, i.e., 1,2,...,n
First, we create a test function with a factor variable:

```{r}
braninFunctionFactor <- function (x) {
   y <- (x[2]  - 5.1/(4 * pi^2) * (x[1] ^2) + 5/pi * x[1]  - 6)^2 +
     10 * (1 - 1/(8 * pi)) * cos(x[1] ) + 10
   if(x[3]==1)
     y <- y +1
   else if(x[3]==2)
     y <- y -1
   return(y)
}
```

Vectorize the test function.

```{r}
objFun <- function(x){apply(x,1,braninFunctionFactor)}
```

Run `spot`.

```{r, spotFac}
set.seed(1)
res <- spot(fun=objFun,lower=c(-5,0,1),upper=c(10,15,3),
            control=list(model=buildKriging,
                         types= c("numeric","numeric","factor"),
                         optimizer=optimLHD))
 res$xbest
 res$ybest
```

### High dimensional problem 

```{r, defineBounds}
n <- 10
a <- rep(0,n)
b <- rep(1,n)
```


First, we consider the default `spot` setting with `buildKriging()`.

```{r, highDimKrig}
tic <- proc.time()[3]
res0 <- spot(x=NULL, funSphere, lower = a, upper = b, 
             control=list(funEvals=30))
toc <- proc.time()[3]
sprintf("value: %f, time: %f",  res0$ybest, toc-tic)
```

Then, we use the `buildGaussianProcess()` model.

```{r, highDimGP}
tic <- proc.time()[3]
res1 <-  spot(x=NULL, funSphere, lower = a, upper = b, 
             control=list(funEvals=30, 
                          model = buildGaussianProcess))
toc <- proc.time()[3]
sprintf("value: %f, time: %f",  res1$ybest, toc-tic)
```



### Run SPOT with logging

```{r}
## run spot without log
res <- spot(fun = funSphere,
            lower=c(0,0),
            upper=c(100,100)
)
## run spot with log
funSphereLog <- function(x){
  cbind(funSphere(x),x)
}
res2 <- spot(fun = funSphereLog,
            lower=c(0,0),
            upper=c(100,100)
)
res$logInfo
res2$logInfo
```

## Hybrid optimization

```{r, hybrid}
res <- spot(fun = funSphere, lower = c(-5,-5),
                upper = c(5,5), 
                control = list(funEvals = 20,
                directOpt = optimNLOPTR,
                directOptControl = list(funEvals = 10)
                ))
str(res)
```



## Handling constraints

* Definitions and constraints

```{r, constraintsSetup, eval = FALSE}
library(babsim.hospital)
n <- 29 
reps <- 2
funEvals <- 3*n 
size <- 2*n
x0 <- matrix(as.numeric(babsim.hospital::getParaSet(5374)[1,-1]),1,)
bounds <- getBounds()
a <- bounds$lower
b <- bounds$upper
g <- function(x) {
      return(rbind(a[1] - x[1], x[1] - b[1], a[2] - x[2], x[2] - b[2], 
                   a[3] - x[3], x[3] - b[3], a[4] - x[4], x[4] - b[4], 
                   a[5] - x[5], x[5] - b[5], a[6] - x[6], x[6] - b[6], 
                   a[7] - x[7], x[7] - b[7], a[8] - x[8], x[8] - b[8], 
                   a[9] - x[9], x[9] - b[9], a[10] - x[10], x[10] - b[10],
                   a[11] - x[11], x[11] - b[11], a[12] - x[12],  x[12] - b[12],
                   a[13] - x[13], x[13] - b[13], a[14] - x[14],  x[14] - b[14],
                   a[15] - x[15], x[15] - b[15], a[16] - x[16],  x[16] - b[16],
                   a[17] - x[17], x[17] - b[17], a[18] - x[18],  x[18] - b[18],
                   a[19] - x[19], x[19] - b[19], a[20] - x[20],  x[20] - b[20],
                   a[21] - x[21], x[21] - b[21], a[22] - x[22],  x[22] - b[22],
                   a[23] - x[23], x[23] - b[23], a[24] - x[24],  x[24] - b[24],
                   a[25] - x[25], x[25] - b[25], a[26] - x[26],  x[26] - b[26],
                   a[27] - x[27], x[27] - b[27], x[15] + x[16] - 1, 
                   x[17] + x[18] + x[19] - 1, x[20] + x[21] - 1, x[23] + x[29] - 1)
      )
  }
```


```{r, constraintsRun, eval = FALSE}
res <- spot(
  x = x0,
  fun = funBaBSimHospital,
  lower = a,
  upper = b,
  verbosity = 0,
  control = list(
    funEvals = 2 * funEvals,
    noise = TRUE,
    designControl = list(# inequalityConstraint = g,
      size = size,
      retries = 1000),
    optimizer = optimNLOPTR,
    optimizerControl = list(
      opts = list(algorithm = "NLOPT_GN_ISRES"),
      eval_g_ineq = g
    ),
    model =  buildKriging,
    plots = FALSE,
    progress = TRUE,
    directOpt = optimNLOPTR,
    directOptControl = list(funEvals = 0),
    eval_g_ineq = g
  )
)
print(res)
````

## GECCO Industrial Challenge 2021

A description of the challenge can be found here:
[GECCO Industrial Challenge 2021](https://www.th-koeln.de/informatik-und-ingenieurwissenschaften/gecco-2021-industrial-challenge-call-for-participation_82086.php)

### Evaluation Using the Docker Container

In order to participate in the GECCO Industrial challenge 2021 you will have to 
download the corresponding Docker image (may need sudo rights to download). 
On your terminal console an evaluation in the BabSim.Hospital
should look something like this (take care, the formatting of the symbols - and ’ can cause this command not to work on your terminal):
```{r, eval=FALSE}
# docker run --rm mrebolle/r-geccoc:Track1 -c 'Rscript objfun.R "6,7,3,3,3,5,3,3,25,17,2,1,0.25,0.05,0.07,0.005,0.07,1e-04,0.08,0.25,0.08,0.5,1e-06,2,1e-06,1e-06,1,2,0.5"'
```

The call of the docker command can be implemtented as follows in R:

```{r, eval =FALSE}
library(SPOT)
# docker run --rm mrebolle/r-geccoc:Track1 -c 'Rscript objfun.R "6,7,3,3,3,5,3,3,25,17,2,1,0.25,0.05,0.07,0.005,0.07,1e-04,0.08,0.25,0.08,0.5,1e-06,2,1e-06,1e-06,1,2,0.5"'
evalFun <- function(candidateSolution){
    evalCommand <- paste0("docker run --rm mrebolle/r-geccoc:Track1 -c ", "'","Rscript objfun.R ")
    parsedCandidate <- paste(candidateSolution, sep=",", collapse = ",")
    return(as.numeric(system(paste0(evalCommand, '"', parsedCandidate, '"', "'"), intern = TRUE)))
}

lower <- c(6,7,3,3,3,5,3,3,25,17,2,1,0.25,0.05,0.07,
           0.005,0.07,1e-04,0.08,0.25,0.08,0.5,1e-06,
           2,1e-06,1e-06,1,2,0.5)

# testing: evalFun(lower)

upper<- c(14,13,7,9,7,9,5,7,35,25,5,7,2,0.15,0.11,0.02,
          0.13,0.002,0.12,0.35,0.12,0.9,0.01,4,1.1,0.0625,
          2,5,0.75)

wFun <- wrapFunction(evalFun)

n <- 29 
reps <- 2
funEvals <- 10*n 
size <- 2*n
x0 <- matrix(as.numeric(babsim.hospital::getParaSet(5374)[1,-1]),1,)
bounds <- getBounds()
a <- bounds$lower
b <- bounds$upper
g <- function(x) {
  return(rbind(a[1] - x[1], x[1] - b[1], a[2] - x[2], x[2] - b[2], 
               a[3] - x[3], x[3] - b[3], a[4] - x[4], x[4] - b[4], 
               a[5] - x[5], x[5] - b[5], a[6] - x[6], x[6] - b[6], 
               a[7] - x[7], x[7] - b[7], a[8] - x[8], x[8] - b[8], 
               a[9] - x[9], x[9] - b[9], a[10] - x[10], x[10] - b[10],
               a[11] - x[11], x[11] - b[11], a[12] - x[12],  x[12] - b[12],
               a[13] - x[13], x[13] - b[13], a[14] - x[14],  x[14] - b[14],
               a[15] - x[15], x[15] - b[15], a[16] - x[16],  x[16] - b[16],
               a[17] - x[17], x[17] - b[17], a[18] - x[18],  x[18] - b[18],
               a[19] - x[19], x[19] - b[19], a[20] - x[20],  x[20] - b[20],
               a[21] - x[21], x[21] - b[21], a[22] - x[22],  x[22] - b[22],
               a[23] - x[23], x[23] - b[23], a[24] - x[24],  x[24] - b[24],
               a[25] - x[25], x[25] - b[25], a[26] - x[26],  x[26] - b[26],
               a[27] - x[27], x[27] - b[27], x[15] + x[16] - 1, 
               x[17] + x[18] + x[19] - 1, x[20] + x[21] - 1, x[23] + x[29] - 1)
  )
}

res <- spot(x = x0,
  fun = wFun,
  lower = a,
  upper = b,
  control = list(
    funEvals = 2 * funEvals,
    noise = TRUE,
    designControl = list(# inequalityConstraint = g,
      size = size,
      retries = 1000),
    optimizer = optimNLOPTR,
    optimizerControl = list(
      opts = list(algorithm = "NLOPT_GN_ISRES"),
      eval_g_ineq = g
    ),
    model =  buildKriging,
    plots = TRUE,
    progress = TRUE,
    directOpt = optimNLOPTR,
    directOptControl = list(funEvals = 0),
    eval_g_ineq = g
  )
)


# 
# res <- spot(NULL,
#             wFun, 
#             lower = lower, 
#             upper = upper, 
#             control = list(funEvals=100,
#                            model=buildKriging, 
#                            optimizer = optimLBFGSB,
#                            designControl=list(size=50)))

```

### Evaluation Using the babsim.hospital R Package


```{r, loadLibsBabsim, eval=FALSE}
library(SPOT)
library(babsim.hospital)
```

```{r, constraintsBabsimSetup, eval = FALSE}
library(babsim.hospital)
n <- 29 
reps <- 2
funEvals <- 3*n 
size <- 2*n
x0 <- matrix(as.numeric(babsim.hospital::getParaSet(5374)[1,-1]),1,)
bounds <- getBounds()
a <- bounds$lower
b <- bounds$upper
g <- function(x) {
      return(rbind(a[1] - x[1], x[1] - b[1], a[2] - x[2], x[2] - b[2], 
                   a[3] - x[3], x[3] - b[3], a[4] - x[4], x[4] - b[4], 
                   a[5] - x[5], x[5] - b[5], a[6] - x[6], x[6] - b[6], 
                   a[7] - x[7], x[7] - b[7], a[8] - x[8], x[8] - b[8], 
                   a[9] - x[9], x[9] - b[9], a[10] - x[10], x[10] - b[10],
                   a[11] - x[11], x[11] - b[11], a[12] - x[12],  x[12] - b[12],
                   a[13] - x[13], x[13] - b[13], a[14] - x[14],  x[14] - b[14],
                   a[15] - x[15], x[15] - b[15], a[16] - x[16],  x[16] - b[16],
                   a[17] - x[17], x[17] - b[17], a[18] - x[18],  x[18] - b[18],
                   a[19] - x[19], x[19] - b[19], a[20] - x[20],  x[20] - b[20],
                   a[21] - x[21], x[21] - b[21], a[22] - x[22],  x[22] - b[22],
                   a[23] - x[23], x[23] - b[23], a[24] - x[24],  x[24] - b[24],
                   a[25] - x[25], x[25] - b[25], a[26] - x[26],  x[26] - b[26],
                   a[27] - x[27], x[27] - b[27], x[15] + x[16] - 1, 
                   x[17] + x[18] + x[19] - 1, x[20] + x[21] - 1, x[23] + x[29] - 1)
      )
  }
```


```{r, constraintsBabsimRun, eval = FALSE}
res <- spot(
  x = x0,
  fun = funBaBSimHospital,
  lower = a,
  upper = b,
  verbosity = 0,
  control = list(
    funEvals = 2 * funEvals,
    noise = TRUE,
    designControl = list(# inequalityConstraint = g,
      size = size,
      retries = 1000),
    optimizer = optimNLOPTR,
    optimizerControl = list(
      opts = list(algorithm = "NLOPT_GN_ISRES"),
      eval_g_ineq = g
    ),
    model =  buildKriging,
    plots = FALSE,
    progress = TRUE,
    directOpt = optimNLOPTR,
    directOptControl = list(funEvals = 0),
    eval_g_ineq = g
  )
)
print(res)
````
